---
title: "p8105_hw2_kxg2002.Rmd"
author: "Katie"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl) 
library(dplyr)
```

## Load in the trash wheel data 

```{r}
trash_wheel_data = 
  read_excel(
    path = "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  select(-homes_powered) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

## Read and clean 2017 and 2018 precip data 

```{r}
precip_data_2017 = 
  read_excel(
    path = "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",  
    sheet = 6, 
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2017") 
  
  
precip_data_2018 = 
  read_excel(
    path = "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = 5, 
    range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>%  
  mutate(year = "2018") 
```

## Combine precip data sets 

```{r}
precip_data = 
  full_join(precip_data_2017, precip_data_2018) %>% 
  mutate(month = month.name[month])
```

## Write about the data 

Looking at the two datasets from 2017 and 2018, we can see that there was a substantial amount of precipitation for both years in May and July. However, there were also differences. For example, in 2018 there was a large amount of precipitation in September compared to 2017, where there was not a lot of precipitation. There were `r nrow(precip_data_2017)` observations in the 2017 precipiation data, `r nrow(precip_data_2018)` observations in the 2018 precipitation data, and `r nrow(trash_wheel_data)` observations from the trash wheel data. Key variables in the precipitation data included `r variable.names(precip_data)`. Key variables in the trash wheel data included `r variable.names(trash_wheel_data)`. The total precipitation in 2018 was `r sum(pull(precip_data_2018, total))`, and the median number of sports balls found in dumpsters in 2017 were `r median(pull(trash_wheel_data, sports_balls))`. 

## Load, read, and clean data in pols-month.csv

```{r}
polls_data = 
  read.csv("./Data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%  
  mutate(month = month.abb[month]) %>% 
  select(year, month, day, prez_gop, prez_dem, everything()) %>% 
  pivot_longer(
    prez_gop:prez_dem,
    names_to = "president", 
    values_to = "status"
    ) %>% 
  select(year, month, president, everything(), -day, -status) 
```

## Load, read, and clean data in snp.csv

```{r}
snp_data = 
  read.csv("./Data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(year = as.integer(year), month = as.integer(month)) %>%  
  mutate(month = month.abb[month]) %>% 
  select(year, month, close, -day)
```

## Load, read, and clean data in unemployment.csv

```{r}
unemployment_data = 
  read_csv("./Data/fivethirtyeight_datasets/unemployment.csv") %>% 
  mutate(Year = as.integer(Year)) %>%  
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment") %>%  
    janitor::clean_names()
```

## Merge data sets 

```{r}
join_polls_and_snp = 
  left_join(polls_data, snp_data)

all_joined_datasets = 
  left_join(join_polls_and_snp, unemployment_data) 
```

## Write about the data 

The polls data set contained information on the number and breakdown of current and past serving Democratic and Republican public officials, specifically governors, senators, and representatives. It also included the party representation of the President. The dimensions of the data set were `r dim(polls_data)`, the names of key variables included `r variable.names(polls_data)`, and the range of years listed were between `r range(pull(polls_data, year))`. 

The snp data set contained information on the closing values of the S&P stock index on the date listed. The dimensions of the data set were `r dim(snp_data)`, the names of key variables included `r variable.names(snp_data)`, and the range of years listed were between `r range(pull(snp_data, year))`. 

The unemployment data set contained information on the percentage of unemployment in the corresponding month and year. The dimensions of the data set were `r dim(unemployment_data)`, the names of key variables included `r variable.names(unemployment_data)`, and the range of years listed were between `r range(pull(unemployment_data, year))`. 

The resulting combined dataset of all three contained information on the number of Democratic and Republican public officials and the offices they held during a specific month and year, the closing values of the S&P stock index during that time frame, and the percentage of unemployment that month and year. 

## Load and tidy Popular_Baby_Names.csv

```{r}
baby_names = 
  read_csv("./Data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender), 
         ethnicity = str_to_lower(ethnicity), 
         childs_first_name = str_to_lower(childs_first_name), 
         ethnicity = recode(ethnicity, 
                            "asian and paci" = "asian and pacific islander", 
                            "black non hisp" = "black non hispanic", 
                            "white non hisp" = "white non hispanic")) %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all= TRUE)
```

## Create a table to show the rank in popularity of the name Olivia over the years 

```{r}
olivia_table = 
  filter(baby_names, childs_first_name == "olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = year_of_birth, 
    values_from = rank) 
```

## Create a table to show the most popular name among male children over time 

```{r}
most_popular_boy_names = 
  filter(baby_names, gender == "male", rank == "1") %>%  
  select(-gender, -count, -rank) %>%
  pivot_wider(
    names_from = year_of_birth, 
    values_from = childs_first_name)
```





